Model Comparisons

### 2015-11-30 models/Model_Comparison.Rmd
Model comparisons with reference to RSA+AmbiguousSemantics.Rmd and Model_additionalAlts.Rmd

# 3.c)
# Emp alts
# Normalized = T
# alpha tune = T
# compare.data = Toggle this with data.partial for e6 or NULL for e11
alphas[11] = tune.alpha(data.full, type = "full", compare.data=data.partial, addMid=F, normalize=T)
m3.c = data.full %>%
  do(run.full(., alpha = alphas[11], addMid=F, normalize=T))
# Match with e6 w/ checks
m3.c.matched = m3.c[matched.items, ]
m3.c.matched$degree = ifelse(m3.c.matched$degree == "hi1", "hi", "low")
m3.c.matched = cbind(m3.c.matched, data.partial$listener.p)
all(m3.c.matched$scale == data.partial$scale & m3.c.matched$degree == data.partial$degree)
colnames(m3.c.matched)[length(colnames(m3.c.matched))] = "e6.listener.p"
# Store output
performance.output[11, ] = c("m3.c", round(cor(m3.c.matched$e6.listener.p, m3.c.matched$pred), 5),
                             round(cor(m3.c.matched$listener.p, m3.c.matched$pred), 5), T)
# Store plot
m3.c.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m3.c.matched,
      main=paste("m3.c\nalpha: ", alphas[11],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[11, 2]),
      ylab="Posterior p(rating | word)") +
      ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 3.d)
# Emp alts
# Normalized = F
# alpha tune = T
# compare.data = Toggle this with data.partial or NULL
alphas[12] = tune.alpha(data.full, type = "full", compare.data=data.partial, addMid=F, normalize=F)
m3.d = data.full %>%
  do(run.full(., alpha = alphas[12], addMid=F, normalize=F))
# Match with e6 w/ checks
m3.d.matched = m3.d[matched.items, ]
m3.d.matched$degree = ifelse(m3.d.matched$degree == "hi1", "hi", "low")
m3.d.matched = cbind(m3.d.matched, data.partial$listener.p)
all(m3.d.matched$scale == data.partial$scale & m3.d.matched$degree == data.partial$degree)
colnames(m3.d.matched)[length(colnames(m3.d.matched))] = "e6.listener.p"
# Store output
performance.output[12, ] = c("m3.d", round(cor(m3.d.matched$e6.listener.p, m3.d.matched$pred), 5),
                             round(cor(m3.d.matched$listener.p, m3.d.matched$pred), 5), F)
# Store plot
m3.d.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m3.d.matched,
      main=paste("m3.d\nalpha: ", alphas[12],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[12, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

@@

# Store plot
m4.c.plot = qplot(stars, listener.p, col=degree, 
      data=m4.c.matched,
      main=paste("m4.c\nalpha: ", alphas[15],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[15, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 4.d)
# Emp alts + extras
# Normalized = F
# alpha tune = T
# compare.data = Toggle this with data.partial or NULL
alphas[16] = tune.alpha(data.full.extras, type = "full", compare.data=data.partial, addMid=T, normalize=F)
m4.d = data.full.extras %>%
  do(run.full(., alpha = alphas[16], addMid=T, normalize=F))
# Match with e6 w/ checks
m4.d.matched = m4.d[matched.items.extras, ]
m4.d.matched$degree = ifelse(m4.d.matched$degree == "hi1", "hi", "low")
m4.d.matched = cbind(m4.d.matched, data.partial$listener.p)
all(m4.d.matched$scale == data.partial$scale & m4.d.matched$degree == data.partial$degree)
colnames(m4.d.matched)[length(colnames(m4.d.matched))] = "e6.listener.p"
# Store output
performance.output[16, ] = c("m4.d", round(cor(m4.d.matched$e6.listener.p, m4.d.matched$pred), 5),
                             round(cor(m4.d.matched$listener.p, m4.d.matched$pred), 5), F)
# Store plot
m4.d.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m4.d.matched,
      main=paste("m4.d\nalpha: ", alphas[16],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[16, 2]),
      ylab="Posterior p(rating | word)") +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)
```

# Plots

@@

Initial thoughts:

We see fairly steady model improvement with the addition of salient alternatives. Let's focus on the normalized family of models (1a, 1c, ... , 4a, 4c) - I explain why this might be a good idea below, but I'm not totally convinced. Probably the most important comparisons to pay attention to is the jump in correlation values when we first introduce empirically derived alternatives (models m3). Comparing m2 models (entailment + generic None) and m3 models (empirically derived alternatives) we see jumps of +0.23 for r values with alpha tuning. See the plots below:
```{r}
m2.c.plot
m3.c.plot
```
Importantly, we also see improvements with the addition of 'fitted' alternatives in m4 see the plots below:
```{r}
m3.c.plot
m4.c.plot
```

Additional notes:

1) In general model performance is greater (correlation) for e11 data. This may be due in part to less extreme literal listener evaluations in e11 (see RSA+AmbiguousSemantics for a more detailed analysis for literal listener experiment differences). Our best model performance occurs with e11 and the model is m4.c, correlation = `r round(performance.outpu[15, 3], 5)`.

2) Non-normalized models seem to be more sensitive to alpha tuning as the number of alternatives increases (see m4.d and m3.d with alpha = 8). I'm inclined to stick with normalized model results because of this - we're interested in whether we see model improvement from the addition of salient alternatives - not from tuning hyperparameters. We tend to see more model improvement with alternatives when we're looking at normalized models (see `m3.c` vs `m4.c`).

3) We may want to explore the the idea of unique alphas for a given scale (would need to change how we've written RSA), for example using scale entropy measures.
```{r overall_perf}
performance.output = cbind(performance.output, alphas)
grid.table(performance.output)

m1.a.plot
m1.b.plot
m1.c.plot
m1.d.plot

@@

m3.a.plot
m3.b.plot
m3.c.plot
m3.d.plot

m4.a.plot
m4.b.plot
m4.c.plot
m4.d.plot
```

>>>>

# 3.c)
# Emp alts
# Normalized = T
# alpha tune = T
# compare.data = Toggle this with data.partial for e6 or NULL for e11
alphas[11] = tune.alpha(data.full, type = "full", compare.data=data.partial, addMid=F, normalize=T)
m3.c = data.full %>%
  do(run.full(., alpha = alphas[11], addMid=F, normalize=T))
# Match with e6 w/ checks
m3.c.matched = m3.c[matched.items, ]
m3.c.matched$degree = ifelse(m3.c.matched$degree == "hi1", "hi", "low")
m3.c.matched = cbind(m3.c.matched, data.partial$listener.p)
all(m3.c.matched$scale == data.partial$scale & m3.c.matched$degree == data.partial$degree)
colnames(m3.c.matched)[length(colnames(m3.c.matched))] = "e6.listener.p"
# Store output
performance.output[11, ] = c("m3.c", round(cor(m3.c.matched$e6.listener.p, m3.c.matched$pred), 5),
                             round(cor(m3.c.matched$listener.p, m3.c.matched$pred), 5), T)

# Store plot 2 - both e6 and e11 data
m3.c.jointD = m3.c.matched %>%
  gather(study, listener.pred, listener.p, e6.listener.p)
m3.c.jointD$study = ifelse(m3.c.jointD$study == "e6.listener.p", "e6", "e11")
m3.c.jointPlot = qplot(stars, listener.pred, col=degree, shape=study,
      data=m3.c.jointD,
      main=paste("m3.c\nalpha: ", alphas[11],
                 "\nNormalized = ", T,
                 ylab="Posterior p(rating | word)")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 3.d)
# Emp alts
# Normalized = F
# alpha tune = T
# compare.data = Toggle this with data.partial or NULL
alphas[12] = tune.alpha(data.full, type = "full", compare.data=data.partial, addMid=F, normalize=F)
m3.d = data.full %>%
  do(run.full(., alpha = alphas[12], addMid=F, normalize=F))
# Match with e6 w/ checks
m3.d.matched = m3.d[matched.items, ]
m3.d.matched$degree = ifelse(m3.d.matched$degree == "hi1", "hi", "low")
m3.d.matched = cbind(m3.d.matched, data.partial$listener.p)
all(m3.d.matched$scale == data.partial$scale & m3.d.matched$degree == data.partial$degree)
colnames(m3.d.matched)[length(colnames(m3.d.matched))] = "e6.listener.p"
# Store output
performance.output[12, ] = c("m3.d", round(cor(m3.d.matched$e6.listener.p, m3.d.matched$pred), 5),
                             round(cor(m3.d.matched$listener.p, m3.d.matched$pred), 5), F)

# Store plot 2 - both e6 and e11 data
m3.d.jointD = m3.d.matched %>%
  gather(study, listener.pred, listener.p, e6.listener.p)
m3.d.jointD$study = ifelse(m3.d.jointD$study == "e6.listener.p", "e6", "e11")
m3.d.jointPlot = qplot(stars, listener.pred, col=degree, shape=study,
      data=m3.d.jointD,
      main=paste("m3.d\nalpha: ", alphas[12],
                 "\nNormalized = ", T,
                 ylab="Posterior p(rating | word)")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

@@

# Store plot 2 - both e6 and e11 data
m4.c.jointD = m4.c.matched %>%
  gather(study, listener.pred, listener.p, e6.listener.p)
m4.c.jointD$study = ifelse(m4.c.jointD$study == "e6.listener.p", "e6", "e11")
m4.c.jointPlot = qplot(stars, listener.pred, col=degree, shape=study,
      data=m4.c.jointD,
      main=paste("m4.c\nalpha: ", alphas[15],
                 "\nNormalized = ", T,
                 ylab="Posterior p(rating | word)")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 4.d)
# Emp alts + extras
# Normalized = F
# alpha tune = T
# compare.data = Toggle this with data.partial or NULL
alphas[16] = tune.alpha(data.full.extras, type = "full", compare.data=data.partial, addMid=T, normalize=F)
m4.d = data.full.extras %>%
  do(run.full(., alpha = alphas[16], addMid=T, normalize=F))
# Match with e6 w/ checks
m4.d.matched = m4.d[matched.items.extras, ]
m4.d.matched$degree = ifelse(m4.d.matched$degree == "hi1", "hi", "low")
m4.d.matched = cbind(m4.d.matched, data.partial$listener.p)
all(m4.d.matched$scale == data.partial$scale & m4.d.matched$degree == data.partial$degree)
colnames(m4.d.matched)[length(colnames(m4.d.matched))] = "e6.listener.p"
# Store output
performance.output[16, ] = c("m4.d", round(cor(m4.d.matched$e6.listener.p, m4.d.matched$pred), 5),
                             round(cor(m4.d.matched$listener.p, m4.d.matched$pred), 5), F)

# Store plot 2 - both e6 and e11 data
m4.d.jointD = m4.d.matched %>%
  gather(study, listener.pred, listener.p, e6.listener.p)
m4.d.jointD$study = ifelse(m4.d.jointD$study == "e6.listener.p", "e6", "e11")
m4.d.jointPlot = qplot(stars, listener.pred, col=degree, shape=study,
      data=m4.d.jointD,
      main=paste("m4.d\nalpha: ", alphas[16],
                 "\nNormalized = ", T,
                 ylab="Posterior p(rating | word)")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)
```

# Discussion

@@

```{r}
performance.output$cor.e6 = as.numeric(performance.output$cor.e6)
performance.output$cor.e11 = as.numeric(performance.output$cor.e11)
performance.output = performance.output %>%
  mutate(avg.corr =
           (cor.e6 + cor.e11) / 2)
performance.output = cbind(performance.output, alphas)
# reorder columns
performance.output = performance.output[, c(1, 2, 3, 5, 4, 6)]
grid.table(performance.output)
```

## Initial thoughts:

### Overall

Across e6 and e11 pragmatic listener judgements we see model fit improve as we add more alternatives. The largest improvement in model fit occurs when we first include the empirically derived salient alternatives (correlations jump +0.23). Have a look at the `m2` vs `m3` correlation for models with pre-RSA normalization below (`c` models):
```{r}
# Tuned, normalized models r
barchart.d = performance.output[c(3, 7, 11, 15), ] %>%
  gather(study, cor, cor.e6, cor.e11)
barchart.d$cor = as.numeric(barchart.d$cor)
ggplot(barchart.d, aes(x=model, y=cor, fill=study)) + 
  geom_bar(stat="identity", position="dodge") +
  ylim(0, 1) + 
  ggtitle("Model fit for e6 and e11\npragmatic listener judgments") +
  geom_text(aes(x=model, y=as.numeric(cor) + 0.025, label=round(cor, 3), col=(study)))
```

Here's another look at the improvement in model fit between normalized `m2` and `m3`.
```{r}
m2.c.plot
m3.c.jointPlot
```

Importantly, we also see improvements with the addition of 'fitted' alternatives in m4 see the plots below:
```{r}
m3.c.jointPlot
m4.c.jointPlot
```

This presents a little bit more of a challenge, however, as we'd like to jointly fit our alpha as well as our alternatives - is this worth delving into? Currently we're just fitting alpha and then finding alternatives.

## Additional notes:

1) In general model performance is greater (correlation) for e11 data. This may be due in part to less extreme literal listener evaluations in e11 see:

https://cdn.rawgit.com/langcog/scalar_implicature/master/models/RSA%2BAmbiguousSemantics.html#pragmatic-listener-studies---judgement-distributional-differences

Our best model performance occurs with e11 and the model is m4.c, correlation = `r as.numeric(performance.output[15, 3])` (see plot below):
```{r}
qplot(stars, listener.p, col=degree, 
      data=m4.c.matched,
      main=paste("m4.c\nalpha: ", alphas[15],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[15, 3]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)
```

Normalizing vs not normalizing before running RSA:

Non-normalized models seem to be more sensitive to alpha tuning as the number of alternatives increases (see m4.d and m3.d with alpha = 8). I'm not sure if there is a theoretical consideration here that RSA may want to address? What does it mean to normalize literal listener values prior to computing the posterior? One repercussion is that during normalization we'll see larger impact from salient alternatives... In `Model_addtionalAlts.Rmd` I'm seeing that higher entropy scales `memorable_unforgettable`, `palatable_delicious` appear to be looking for more alternatives (correlation improves when I add wacky distributions such as [0, 0.1, 0.9, 0.1, 0.9] which suggests that we need possibly *two* more alternatives, not just one). This seems intuitive in the higher entropy setting... Can we interpret this kind of finding (at least intuitively) as the model telling us it wants *more* alternatives? See

https://cdn.rawgit.com/langcog/scalar_implicature/master/models/Model_additionalAlts.html#match-items-between-studies-full-and-partial

# All plots with e6 prag judgements
```{r overall_perf}
m1.a.plot
m1.b.plot
m1.c.plot
m1.d.plot

@@

m3.a.plot
m3.b.plot
m3.c.jointPlot
m3.d.jointPlot

m4.a.plot
m4.b.plot
m4.c.jointPlot
m4.d.jointPlot
```
